Looking for a Signal in the Noise: Revisiting Obesity and the microbiome

### 2016-03-17 Sze_ObesityMicrobiomeMetaNoMetaHit_PeerJ_2016.Rmd
Readability and Common OTU Class

Next we investigated how well the OTUs in each study were able to classify obese and non-obese individuals using the random forest algorithm. We found that the average Out Of Bag (OOB) Area Under the Curve (AUC) was 0.7553 +/- 0.0855. We found that the total range for the total number of variables used for the classification ranged from 3 (Ross and Escobar) to 40 (Zupancic). The total number of trees and node size was standardized to 1000 and 20 respectively for each data set. One aspect we found surprising was that there was very little overlap in the variables used by the random forest algorithm for the classification of obese and non-obese individuals by study. The three most common variables was for an OTU that could be classified to the family Ruminococcaceae (7/9 studies), an OTU that could be classified to the family Lachnospiraceae (5/9 studies), and the Shannon diversity metric (4/9 studies).

>>>>

Next we investigated how well the OTUs in each study were able to classify obese and non-obese individuals using the random forest algorithm. We found that the average Out Of Bag (OOB) Area Under the Curve (AUC) was `r format(round(mean(AUCRFData$OOBAUCAll), 2), nsmall=2)` +/- `r format(round(sd(AUCRFData$OOBAUCAll), 2), nsmall=2)`. We found that the total range for the total number of variables used for the classification ranged from 3 (Ross and Escobar) to 40 (Zupancic). The total number of trees and node size was standardized to 1000 and 20 respectively for each data set. One aspect we found surprising was that there was very little overlap in the variables used by the random forest algorithm for the classification of obese and non-obese individuals by study. The three most common variables was for an OTU that could be classified to the family Ruminococcaceae (7/8 studies), an OTU that could be classified to the family Lachnospiraceae (6/8 studies), Clostridiales (6/8) and Shannon diversity metric (5/8 studies).
